Rust Bug Fix Patterns

### 2022-09-19 methodology.tex
comments

\subsubsection{Path Linearization}

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{linearization.png}
\label{f2}
\caption{Path Linearization}
\end{figure}

The output of dictdiff is a list of diffs, where each diff is constituted of three parts. The first part specifies whether the modification is of type 'add', 'remove', 'change'. The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change and includes all the involved terminals and non terminals in the change. Our tool looks for changes within each root-level scope. That is, our assumption is that a change pattern occurs within one root-level scope. Our perspective does not cover the patterns that involve changes in multiple functions or multiple files. However, if the change happens in multiple lines within one scope, whether they are continuously one after another or separate from each other, our tool would be able to detect it.

Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change. Through a Depth First Search we can collect all of these paths. We store only the non terminals that exist in the Syn token types list. Also certain keywords are not considered as non terminals in the syntax tree (e.g. clone, Box), but as they have a special meaning the language, we also store them in the path. The outcome of the DFS search is a set of linearized paths.

Fig \ref{f2} shows a simple change and its respective linearized paths. In the change, the second argument sends a borrow of the variable as opposed to sending the ownership entirely, and in the third argument, instead of sending the head into the function scope, we send a clone of it. A simplification of the derived diff of the two code revisions is shown in the figure. A DFS search on this tree would yield three different paths shown in three different colors. The path linearizes the hierarchy of non terminals involved in that specific path which represents one contributing factor to the change. We found it crucial to record the order of all non terminals involved in a path as it gives us freedom to control the degree of order-sensitiveness in our datapoint encoding.

The next question is how to transform these paths to a fixed size datapoint. First step, we need to define a fixed set of columns. The number of columns would depend on the amount of information we want to encode in the datapoints. The naivest approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such encoding, for instance, there is no difference between two nested if statements and two if statements beside each other. Theoritically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions. Apart from that, not all combinations adheres to Rust syntax. A reasonable work around would be to categorize the non terminals and set the dimensions as all possible combinations of these categories (Similar to JS paper, ref). This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations. However, after data collections we would still encounter a sparse dataset.

In our methodology, similar to the first approach, we define the columns of our dataset, as the set of token types we collected from Syn. However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps. Firstly, we record the number of occurrences of token types within the paths at each scope. The reason behind this decision is because different order of program elements yields a different number of occurrences of underlying token types. Secondly, we semi-automatically design a weighting scheme to give more value to certain non terminal we think are more important in Rust.

@@

Our main objective is that our datapoints manifest the change with high level of precision. For instance, the purple path in Fig \ref{f2} can be described as a change in local variable declaration, a change in a function call, calling a method of one of the function arguments, or calling clone() of one of the function arguments. Obviously, the last description is the most favorable one, as it offers the most accurate description. The whole idea behind our weighting scheme is to be able to identify the change accurately.

A simple heuristic is that the closer we get to the leaves in the ASTDiff, the more important token types get. Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (Similar to ItemFn, Local, ExprCall in Fig \ref{f2}). Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of token types, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline. We recorded the number of occurrences of each token type in total. Then we gave the value of the inverse of the number of occurrences to each token. The intuition is that the higher the number of occurrences of a token type, the closer it is usually seen to the root. Hence, it probably does not provide us with the most accurate description of the change. In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the token types related to borrow checker. That is why we characterize our weighting scheme as semi-automatically designed.

>>>>

\subsubsection{Path Extraction}

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{linearization.png}
\caption{\label{fig:extraction}Path Extraction}
\end{figure}

% again try to remove redundancy
The output of dictdiff is a list of diffs, where each diff comprises three parts. The first part specifies whether the modification is of type 'add', 'remove', 'change'. The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change (terminals and nonterminals). Our tool looks for changes within each root-level scope. That is, our assumption is that a change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, if the change happens in multiple lines within one scope, whether they are continuously one after another or separate from each other, our tool would be able to detect it.
% simplify that last sentence there
% provide an example and then transform it into fig:extraction.

% what is the Syn token types list?
Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one contributing element to the change. Through a Depth First Search we can collect all of these paths. We store only the non terminals that exist in the Syn token types list. Also certain keywords are not considered as non terminals in the syntax tree (e.g. clone, Box), but as they have a special meaning the language, we also store them in the path. The outcome of the DFS search is a set of paths.

% reorganize this when you put the example up there.
Figure~\ref{fig:extraction} shows a simple change and its respective paths. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it. A DFS search on this tree would yield three different paths, which we show in red, green, and purple. The path represents the sequence of involved non terminals, which represents one contributing factor to the change (along with the context, which we do not record). We found it crucial to record the order of all non terminals involved in a path as it gives us freedom to control the degree of order-sensitiveness in our datapoint encoding.
% can use fewer words in that last sentence too

% OK so if you just have all observed nonterminals then you have a fixed size, but it's too big.
The next question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. The number of columns would depend on the amount of information we want to encode in the datapoints. The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other. Theoretically, to reach full order-sensitiveness, we would need encode all possible combinations of token types as our dimensions, which yields an astronomical number of dimensions. Apart from that, not all combinations adhere to Rust syntax. A reasonable work around would be to reduce from the full set of nonterminals to a smaller set of categories, and then to set the dimension as all possible combinations of these categories (Similar to JS paper, ref). This would reduce the number of dimensions and provide more reasonable and syntactically-correct combinations. However, this would still yield a sparse dataset. 
% give examples of categories
% can we provide orders of magnitude rather than "astronomical"?

% We make a novel contribution here. 
In our methodology, similar to the first approach, we define the columns of our dataset, as the set of token types we collected from Syn. However, to achieve an acceptable level of order-sensitiveness, we carried out two auxiliary steps. First, we record the number of occurrences of token types within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying token types. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs.

@@

% consider: "datapoints accurately describe the most important change." 
% put the most important change first
Our main objective is that our datapoints manifest the change with high level of precision. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description offers the most accurate description. The whole idea behind our weighting scheme is to be able to identify the change accurately.

% I need to reread this key paragraph. - PL
A simple heuristic is that, the closer we get to the leaves in the ASTDiff, the more important token types get. Another observation is that the elements closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). Exploiting the latter observation, to obtain a basic knowledge about the number of repetitions of token types, we mined the last 20 bug related commits of the last 20 most starred rust projects on GitHub, and ran them through of pipeline. We recorded the number of occurrences of each token type in total. Then we gave the value of the inverse of the number of occurrences to each token. The intuition is that the higher the number of occurrences of a token type, the closer it is usually seen to the root. Hence, it probably does not provide us with the most accurate description of the change. In addition to that, as we wanted to make sure that our tool captures the patterns related to borrow checker, we manually raised the weights of the token types related to borrow checker. That is why we characterize our weighting scheme as semi-automatically designed.
